Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-04-18 Manuscript/manuscript.Rmd
manuscript edits

Further improvement in performance was found when models were implemented on both randomly undersampled (RUS) and SMOTE resampled data, for all 3 predictor types (Figure 5). Although the differences in performance between the two techniques were marginal, SMOTE resampled data was found to yield the best F1-Scores and MCCs among all classes of models developed and compared in the ensemble framework. As with ROS, distance type predictors yeilded the highest F1-Score and MCC values across resolutions, when applied to either RUS or SMOTE. SMOTE re-sampled data, with the inclusion of distance-type predictors was also found to be the best set of characteristics when applied to the K562 cell line (Supplementary Figure 10). These results imply that a best performing model, one that is able to best discriminate between genomic bins that are near versus far away from TAD boundaries, is one that incorporates both random under- and over-sampling, as SMOTE is known to do. Likewise, it appears that distance-type predictors provided the most informative spatial representation of epigenomic annotations associated with TAD boundaries.

@@

>>>>

Further improvement in performance was found when models were implemented on both randomly undersampled (RUS) and SMOTE resampled data, for all 3 predictor types (Figure 5). Although the differences in performance between the two techniques were marginal, SMOTE resampled data was found to yield the best F1-Scores and MCCs among all classes of models developed and compared in the ensemble framework. As with ROS, distance type predictors yeilded the highest F1-Score and MCC values across resolutions, when applied to either RUS or SMOTE. SMOTE re-sampled data, with the inclusion of distance-type predictors was also found to be the best set of characteristics when applied to the K562 cell line (Supplementary Figure 10).

@@

These results imply that a best performing model, one that is able to best discriminate between genomic bins that are near versus far away from TAD boundaries, is one that incorporates both random under- and over-sampling, as SMOTE is known to do.  Additionally, when we performed and evaluated the variable importances of random forest models built on one genomic annotation at a time (including the 3 predictor types each), we found that the distance-type consistently out-ranked the OC and OP type predictors in terms of predictive importance, for a majority of the cell line specific annotations, across resolution and cell line (Supplementary Tables 4 & 5). Therefore, it appears that distance-type predictors provided the most informative spatial representation of epigenomic annotations associated with TAD boundaries.
